---
title: '02 Cleaning - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "02-cleaning.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [RNAseq-preprocess](https://github.com/eturkes/RNAseq-preprocess).*

In this document we prepare the gene count matrix for downstream analysis.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[6]``.

```{r}
# Copyright 2020 Emir Turkes, UK DRI at UCL
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
#     http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

# This section should be checked per document.
# Blank variables should be set to NULL.
# --------------------------------------------
options(stringsAsFactors = FALSE)
packages <- c("conflicted", "biomaRt")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
analysis_no <- 2
data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
assets_dir <- file.path(getwd(), "..", "..", "assets", data_name) # Backed up caches and data.

tmp_dir <- file.path(getwd(), "..", "..", "tmp", data_name)
if (!dir.exists(tmp_dir)) {
  dir.create(tmp_dir, recursive = TRUE)
}

# Unique cache directory for each analysis number.
# ------------------------------------------------
if (!dir.exists(file.path(assets_dir, "cache", paste0("0", analysis_no)))) {
  dir.create(file.path(assets_dir, "cache", paste0("0", analysis_no)), recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7)
# ------------------------------------------------
# ----------------------------------------------------------
```

# Cleaning of Count Matrix

```{r}
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "counts_cleaned.rds")
rds2 <- file.path(assets_dir, "cache", paste0("0", analysis_no), "gene_anno.rds")
if (file.exists(rds)) {
  counts_cleaned <- readRDS(rds)
  gene_anno <- readRDS(rds2)
} else {
  file.path(assets_dir, "cache", paste0("0", analysis_no - 1), "feature_counts.rds")
  counts_cleaned <- feature_counts$counts

  colnames(counts_cleaned) <- sub(".FCH.*", "", colnames(counts_cleaned))
  rownames(counts_cleaned) <- sub("\\..*", "", rownames(counts_cleaned))

  mart <- useEnsembl("ensembl", "mmusculus_gene_ensembl")
  attributes <- c("external_gene_name", "ensembl_gene_id")
  gene_anno <- getBM(attributes, "ensembl_gene_id", rownames(counts_cleaned), mart)

  # This is an ugly solution that should be looped until there are no duplicates left, but for now
  # it works.
  # ----------------------------------------------------------------------------------------------
  dup <- gene_anno[duplicated(gene_anno$external_gene_name), ]
  if (nrow(dup) > 0) {
    for (i in 1:dim(dup)[1]) {
      for (j in 1:dim(gene_anno)[1]) {
        if (dup$ensembl_gene_id[i] == gene_anno$ensembl_gene_id[j]) {
          gene_anno$external_gene_name[j] <- paste0(gene_anno$external_gene_name[j], "-alternative")
        }
      }
    }
  }
  dup <- gene_anno[duplicated(gene_anno$external_gene_name), ]
  if (nrow(dup) > 0) {
    for (i in 1:dim(dup)[1]) {
      for (j in 1:dim(gene_anno)[1]) {
        if (dup$ensembl_gene_id[i] == gene_anno$ensembl_gene_id[j]) {
          gene_anno$external_gene_name[j] <- paste0(gene_anno$external_gene_name[j], "2")
        }
      }
    }
  }
  dup <- gene_anno[duplicated(gene_anno$external_gene_name), ]
  if (nrow(dup) > 0) {
    for (i in 1:dim(dup)[1]) {
      for (j in 1:dim(gene_anno)[1]) {
        if (dup$ensembl_gene_id[i] == gene_anno$ensembl_gene_id[j]) {
          gene_anno$external_gene_name[j] <- sub(
            "[[:digit:]]$", "3", gene_anno$external_gene_name[j]
          )
        }
      }
    }
  }
  # ----------------------------------------------------------------------------------------------

  counts_cleaned <- counts_cleaned[rownames(counts_cleaned) %in% gene_anno$ensembl_gene_id, ]
  gene_anno <- gene_anno[order(match(gene_anno$ensembl_gene_id, rownames(counts_cleaned))), ]
  rownames(counts_cleaned) <- gene_anno$external_gene_name

  saveRDS(counts_cleaned, rds)
  saveRDS(gene_anno, rds2)
}
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
